---
title: "Bat date palm visits EDA"
author: "Kelly Endres"
date: "1/24/2022"
output: html_document
---

Clif - when double checking the data I found two trees that have names but where tree type is na. We had originally proceeded with dropping these since we decided that they were flagged but no observations occurred. These are still dropped in the analysis, but we may need to check that these were actually date palm trees, since it would change the number of villages that had date palm sap collection occurring. (Trees 1051.ICT1 and 1051.ICT2)
- line 54

- another tree (2084.ICF1) had nas for all of its values, was this tree also not observed? This would be the same issue as those above, dropped in the analysis thus far.
- line 74 

Bat camera vists - date palm  data

This R Markdown covers exploratory data analysis for bat camera visits at located at date palm trees. Date palm trees were observed at villages where date palm sap collection was reported. 

This file covers initial data examination and additional dataset formation for the three data streams; date palm camera trap observations, fruit tree camera trap observations, and community survey data.

```{r load packages, message = FALSE}
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(naniar)
```

#### Initial data observations

```{r import data}
date_palm_data <- readRDS(here("DATA","date_palm_visit_data.RDS"))
fruit_survey <- fruit_survey <- readRDS(here("DATA", "community_resident_fruit_survey.RDS"))
```

```{r Examine data structure}
# Examine data structure
str(date_palm_data) 

head(date_palm_data)
```

```{r}
# Examine values in main columns used for analysis 
table(is.na(date_palm_data$Village_ID))

table(is.na(date_palm_data$Tree_number))

unique(date_palm_data$Tree_type)
  # different spellings of date palm
  # one column with date as tree type
table(is.na(date_palm_data$Tree_type))
  # if tree type in na, tree was assigned but not observed, these are dropped 
date_palm_data<-date_palm_data[!(is.na(date_palm_data$Tree_number)),]
date_palm_data[which(date_palm_data$Tree_type == "05.01.2013"), ] 
  # replace date of observation
date_palm_data$Date_of_observation <- replace(date_palm_data$Date_of_observation, date_palm_data$Tree_type == "05.01.2013", "05.01.2013")
date_palm_data$Day <- replace(date_palm_data$Day, date_palm_data$Tree_type == "05.01.2013", 5)
date_palm_data$Month <- replace(date_palm_data$Month, date_palm_data$Tree_type == "05.01.2013", 1)
date_palm_data$Year <- replace(date_palm_data$Year, date_palm_data$Tree_type == "05.01.2013", 2013)
date_palm_data[which(date_palm_data$Tree_type == "05.01.2013"), ] 

  # fix tree type spelling 
date_palm_data <- date_palm_data %>% 
  select(-Tree_type) %>% 
  mutate(Tree_type = "Date palm")
unique(date_palm_data$Tree_type)

table(is.na(date_palm_data$DurContT))
  # one with na values
date_palm_data[which(is.na(date_palm_data$DurContT)), ] 
  # na for all measured visit variables
date_palm_data[which(date_palm_data$Village_ID == 2084), ] 
  # drop tree
date_palm_data<-date_palm_data[!(date_palm_data$Tree_number == "2084.ICF1"),]

table(is.na(fruit_data$Date_of_observation))
unique(fruit_data$Day)
unique(fruit_data$Month)
unique(fruit_data$Year)
  # some with NA for year 
date_palm_data[which(is.na(date_palm_data$Year)), ] 
  # all NAs are the same tree, date was entered incorrectly 

date_palm_data$Year <- replace(date_palm_data$Year, date_palm_data$Tree_number == "1065.ICT1" & date_palm_data$Date_of_observation == "2012.12.20", 2021)
date_palm_data$Month <- replace(date_palm_data$Month, date_palm_data$Tree_number == "1065.ICT1" & date_palm_data$Date_of_observation == "2012.12.20", 12)
date_palm_data$Day <- replace(date_palm_data$Day, date_palm_data$Tree_number == "1065.ICT1" & date_palm_data$Date_of_observation == "2012.12.20", 20)
date_palm_data[which(date_palm_data$Tree_number == "1065.ICT1"), ] 


table(is.na(date_palm_data$DurStayT))
unique(date_palm_data$P_NP)
unique(date_palm_data$Number_of_visits)
unique(date_palm_data$Number_of_stays)
unique(date_palm_data$Number_of_contaminations)
```

Dataset includes 5292 rows of information across 27 variables

```{r add village type information}
# Combine with information on village type

# Dataframe with village type information
survey_villages_1 <- fruit_survey %>% 
                        select(correct_villid,case_control_group) #only keep village type information
survey_villages_1 <- survey_villages_1 %>% 
                        rename(Village_ID = correct_villid)
survey_villages_1 <- transform(survey_villages_1, Village_ID = as.numeric(Village_ID))
survey_villages_1 <- survey_villages_1[!duplicated(survey_villages_1), ] #drop duplicate observations
survey_villages_1 <- na.omit(survey_villages_1) #drop NA values
survey_villages_1 <- survey_villages_1 %>% 
  mutate("dup" = duplicated(survey_villages_1$Village_ID))

date_palm_vil <- transform(date_palm_data, Village_ID = as.numeric(Village_ID))
date_palm_vil <-
  left_join(date_palm_data,
            survey_villages_1 %>% dplyr::select(Village_ID, case_control_group))

# recode bat types
date_palm_vil <- date_palm_vil %>% 
  mutate(recode_P_NP = case_when(Number_of_visits == 0 ~ "",
                                 P_NP == 0 ~ "Non Pteropus", 
                                 P_NP == 1 ~ "Pteropus",
                                 TRUE ~ "Unidentified")) %>% 
  replace_with_na(replace = list(recode_P_NP = ""))

saveRDS(date_palm_vil, file= here("Data", "date_palm_with_villages.RDS"))

```