---
title: "Community Resident Fruit Survey EDA"
author: "Kelly Endres"
date: "10/25/2022"
output: html_document
---

Rmarkdown file examining bat, human, and domestic animal fruit consumption behaviors in Bangladesh. Compares behaviors between Nipah virus case and control villages. 

This preliminary data analysis reviews the community resident survey given to the selected baris in each village. The survey assesses fruit consumption, related habits, and fruit trees. Thirty-one different types of fruits/fruit trees are included in the survey. 

Many questions address consumption of dropped fruits, which is used as a proxy for fruits potentially contaminated by bats. For example, if bats often visit a type of tree, and that tree's fruit is eaten off the ground, its possible that contaminated fruit is being consumed. 

Overview:  
1. Initial data observations  
2. Fruit consumption tables, by village type and tree type  
3. Fruit consumption graphs  
4. Additional fruit consumption questions  
5. Other  

### 1. Initial data observations

The following packages are used in analysis:
```{r load packages, message=FALSE}
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
library(haven)
library(knitr)

options(scipen = 0)
```

Import cleaned dataset of community resident surveys. 
```{r import data}
fruit_survey <- readRDS(here("Data","community_resident_fruit_survey.RDS"))
```

Examine data structure

```{r}
head(fruit_survey)

# number of villages
### I switched from `q1_6` to `correct_villid`.
### The latter cleared up some of the issues with incorrect village IDs, e.g., "10" or "10061"
table(is.na(fruit_survey$correct_villid))
n_distinct(fruit_survey$correct_villid)
  # 206 villages

table(is.na(fruit_survey$dataid))
n_distinct(fruit_survey$dataid)
  # 5056 respondents


```

```{r}

  # Clif - why are there 3 observations with no data?
  ### Not sure why. They are blank in the original Stata file. We could ask Emily.
  ### I suspect they just planned on visiting that village but didn't go.
  fruit_survey[fruit_survey$correct_villid=="", ]
```




  





    